---
title: "Syria: Topic Modeling, Classification and Territorial Control"
author: "Michael L. Davies"
subtitle: ""
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Data Preprocessing


```{r}
library(tidyverse)
theme_set(theme_minimal())
```

## Read the data

```{r}
df_raw <- read_csv("df_acled_syr_2017_2021.csv")  %>% 
  rename("id" = "event_id_cnty")
```



## Display Data Table

Data filtered for reporting on territorial change

```{r}
df_territory <- df_raw %>% 
  filter(sub_event_type %in% c("Government regains territory", "Non-state actor overtakes territory")) 

df_territory %>% count(sub_event_type, sort = T)
```


```{r}
library(DT)
```

```{r}
df_sub <- df_territory %>% 
  dplyr::select(event_date, actor1, admin1, event_type, sub_event_type)

datatable(df_sub, options = list(pagelength = 5, scrollX = "400px"))
```

Look at a sample of the text

```{r}
df_territory %>% 
  sample_n(2) %>% 
  pull(notes)
```


## Pre-process the data

- tokenize
- remove stop words

```{r}
# https://www.tidytextmining.com/nasa.html
library(tidytext)

my_stop_words <- c(stop_words$word, "al", "fatalities", "clashes", "ar",
                   "reported", "injuries", "coded")

tidy_territory <- df_territory %>%
  mutate(line = row_number()) %>% #annotate a line where each word comes from
  unnest_tokens(word, notes) %>% #creates one word per row
  #anti_join(stop_words) %>% 
  filter(!word %in% my_stop_words,  # remove an extra word
         str_detect(word, "[a-z]")) # keep only words
```


## Initial simple exploration
What are the most common words? 

```{r}
tidy_territory %>%
  group_by(sub_event_type) %>% 
  count(word, sort = TRUE) %>% 
  ungroup()
```


Word co-ocurrences and correlations

- Networks of words


```{r}
library(widyr)

title_word_pairs <- tidy_territory %>% 
  group_by(sub_event_type) %>% 
  pairwise_count(word, id, sort = TRUE, upper = FALSE) %>% 
  ungroup()

title_word_pairs
```


## Plot networks of co-occurring words 

- These are the pairs of words that occur together most often when the non-state actor overtakes territory.

```{r fig.width=10}
library(ggplot2)
library(igraph)
library(ggraph)

set.seed(1234)
title_word_pairs %>%
  filter(sub_event_type == "Non-state actor overtakes territory") %>% 
  dplyr::select(item1, item2, n) %>% 
  filter(n >= 150) %>%
  dplyr::select(item1, item2, n) %>% 
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n), edge_colour = "cyan4") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, 
                 point.padding = unit(0.2, "lines")) +
  theme_void()
```

These are the pairs of words that occur together most often when the Government regains territory.

```{r}
set.seed(1234)
title_word_pairs %>%
  filter(sub_event_type == "Government regains territory") %>% 
  dplyr::select(item1, item2, n) %>% 
  filter(n >= 150) %>%
  dplyr::select(item1, item2, n) %>% 
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n), edge_colour = "darkred") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, 
                 point.padding = unit(0.2, "lines")) +
  theme_void()
```


Sorted by tf-idf

```{r}
territory_tf_idf <- tidy_territory %>% 
  count(id, word, sort = TRUE) %>%
  ungroup() %>%
  bind_tf_idf(word, id, n)

territory_tf_idf %>% 
  arrange(-tf_idf)
```

```{r}
territory_tf_idf <- full_join(territory_tf_idf, tidy_territory, by = "id")
```

```{r eval=FALSE}
territory_tf_idf %>% 
  filter(!near(tf, 1)) %>%
  filter(sub_event_type %in% c("Government regains territory", "Nonstate group overtakes territory")) %>%
  arrange(desc(tf_idf)) %>%
  group_by(sub_event_type) %>%
  distinct(word, sub_event_type, .keep_all = TRUE) %>%
  slice_max(tf_idf, n = 15, with_ties = FALSE) %>% 
  ungroup() %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>%
  ggplot(aes(tf_idf, word, fill = keyword)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sub_event_type, ncol = 2, scales = "free") +
  labs(title = "Highest tf-idf words",
       x = "tf-idf", y = NULL)
```


Latent Dirichlet Allocation with the topicmodels package


```{r}
# https://juliasilge.github.io/tidytext/articles/topic_modeling.html
# 
# Right now this data frame is in a tidy form, with one-term-per-document-per-row. However, the topicmodels package requires a DocumentTermMatrix (from the tm package). As described in this vignette, we can cast a one-token-per-row table into a DocumentTermMatrix with tidytext’s cast_dtm:

territory_dtm <- tidy_territory %>%
  count(sub_event_type, word, sort = TRUE) %>% 
  cast_dtm(sub_event_type, word, n)

```

```{r}
# In this case, I'm setting it k= 6; in practice we may need to try a few different values of k.
library(topicmodels)
territory_lda <- LDA(territory_dtm, k = 6, control = list(seed = 1234))
territory_lda
```


```{r}
territory_lda_td <- tidy(territory_lda)
```


Find the top 5 terms within each topic:

```{r}
top_terms <- territory_lda_td %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)
```


```{r fig.width=10}
theme_set(theme_bw())

top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  mutate(topic = paste0("Topic ", topic)) %>% 
  ggplot(aes(term, beta, fill = topic)) +
  geom_histogram(stat = "identity", alpha = 0.8) +
  scale_x_reordered() +
  scale_y_continuous(limits = c(0,NA), expand = c(0, 0)) +
  theme(legend.position = "None") +
  coord_flip()+
  facet_wrap(~ topic, scales = "free_y", ncol = 3)
```

Find co-occurances with the word "islamic" as most of the violent groups are accused of being "islamic" in social media.

```{r}
title_word_pairs %>% 
  filter(item1 == "islamic" | item2 == "islamic")
```

## Now looking at bi-grams

```{r}
tidy_bigram <- df_territory %>%
  mutate(line = row_number()) %>% #annotate a line where each word comes from
  unnest_tokens(bigram, notes, token = "ngrams", n = 2) #creates one bigram per row
#anti_join(stop_words) %>% 
#filter(str_detect(word, "[a-z]")) # keep only words

tidy_bigram %>%
  count(bigram, sort = TRUE)

library(tidyr)

my_stop_words <- c(stop_words$word, "al", "coded", "of") #"2020", "2019", "2018", "2017",


bigrams_separated <- tidy_bigram %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% my_stop_words, 
         str_detect(word1, "[a-z]")) %>%
  filter(!word2 %in% my_stop_words, 
         str_detect(word2, "[a-z]"))
#str_detect(word1, [0-9])

# new bigram counts:
bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)

bigrams_united <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ") #Convenience function to paste together multiple columns into one.

bigrams_united %>% 
  filter(str_detect(bigram, "islamic")) %>% 
  count(bigram, sort = TRUE) %>% 
  mutate(bigram = fct_reorder(bigram, n)) %>% 
  head(20) %>% 
  ggplot(aes(bigram, n)) +
  geom_col() +
  coord_flip() +
  #facet_wrap(~ sub_event_type) +
  labs(title = "Common 'Bigrams' with 'Islamic'")

# tidy_bigram %>% 
#   select(sub_event_type, bigram) %>% 
#   filter(str_detect(bigram, "islamic")) %>% 
#   count(sub_event_type)
```

```{r}
bigrams_united %>% 
  filter(str_detect(bigram, "control")) %>% 
  count(bigram, sort = TRUE) %>% 
  mutate(bigram = fct_reorder(bigram, n)) %>% 
  head(20) %>% 
  ggplot(aes(bigram, n)) +
  geom_col() +
  coord_flip() +
  #facet_wrap(~ sub_event_type) +
  labs(title = "Common 'Bigrams' with 'control'")
```

```{r}
bigrams_united %>% 
  count(sub_event_type, bigram, sort = TRUE) %>% 
  mutate(bigram = fct_reorder(bigram, n)) %>% 
  head(20) %>% 
  ggplot(aes(bigram, n)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~sub_event_type, scales = "free") +
  #facet_wrap(~ sub_event_type) +
  labs(title = "Common 'Bigrams' with 'control'")
```


```{r fig.height=8}
bigrams_united %>% 
  count(admin1, sub_event_type, bigram, sort = TRUE) %>% 
  mutate(bigram = fct_reorder(bigram, n)) %>% 
  head(20) %>% 
  ggplot(aes(bigram, n, fill=sub_event_type)) +
  geom_col(position = "dodge") +
  labs(fill = "",
       x = "") +
  coord_flip() +
  facet_wrap(~admin1, scales = "free") +
  theme(legend.position = "bottom")
```




# Topic model

From a different tutorial - possibly text_modeling_tutorial

```{r}
library(quanteda) #for a document frequency matrix (a special case of the doc term matrix)
library(stm)

tidy_text <- tidy_territory

text_dfm <- tidy_text %>%
  count(sub_event_type, word, sort = TRUE) %>%
  cast_dfm(sub_event_type, word, n)

text_sparse <- tidy_text %>%
  count(sub_event_type, word, sort = TRUE) %>%
  cast_sparse(sub_event_type, word, n)
```


```{r}
topic_model <- stm(text_dfm, K = 2, #looking for 2 topics (judgment call?)
                   verbose = FALSE, init.type = "Spectral")
```


```{r}
td_beta <- tidy(topic_model)

td_beta %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  mutate(topic = paste0("Topic ", topic),
         term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free_y") +
  coord_flip() +
  scale_x_reordered() +
  labs(x = NULL, y = expression(beta),
       title = "Highest word probabilities for each topic",
       subtitle = "Different words are associated with different topics")
```


Now, the probability that each document is generated from each topic.

Honestly, I don't really understand this approach.

```{r}
td_gamma <- tidy(topic_model, matrix = "gamma",                    
                 document_names = rownames(text_dfm))

ggplot(td_gamma, aes(gamma, fill = as.factor(topic))) +
  geom_histogram(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~ topic, ncol = 3) +
  labs(title = "Distribution of document probabilities for each topic",
       y = "Number of ?", x = expression(gamma))
```


# Log Reg - bootstrapped

```{r}
# https://juliasilge.com/blog/animal-crossing/

library(tidymodels)
set.seed(123)
model_split <- initial_split(df_territory, strata = sub_event_type)
model_train <- training(model_split)
model_test <- testing(model_split)
```

Preprocess the data for modeling. 

```{r}
library(textrecipes)

model_rec <- recipe(sub_event_type ~ notes + admin2 + fatalities + event_date, data = model_train) %>%
  step_date(event_date, features = c("month", "dow"), role = "dates") %>%
  step_rm(event_date) %>%
  step_dummy(has_role("dates")) %>%
  step_tokenize(notes) %>%
  step_stopwords(notes) %>%
  step_ngram(notes, num_tokens = 2, min_num_tokens = 1) %>%
  step_tokenfilter(notes, max_tokens = 500) %>%
  step_tfidf(notes) %>% 
  step_dummy(admin2) %>% 
  step_normalize(all_predictors())

model_prep <- prep(model_rec)

model_prep %>% juice()
```

Using lasso out of interest.

```{r}
lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lasso_wf <- workflow() %>%
  add_recipe(model_rec) %>%
  add_model(lasso_spec)

lasso_wf
```

# Tune model parameters

Set of possible regularization parameters to try.

```{r}
lambda_grid <- grid_regular(penalty(), levels = 40)

#Next, we need a set of resampled data to fit and evaluate all these models.

set.seed(234)
model_folds <- bootstraps(model_train, strata = sub_event_type)
model_folds
```

```{r}
doParallel::registerDoParallel()

set.seed(2020)
lasso_grid <- tune_grid(
  lasso_wf,
  resamples = model_folds,
  grid = lambda_grid,
  metrics = metric_set(roc_auc, ppv, npv)
)
```


```{r}
# Once we have our tuning results, we can examine them in detail.
lasso_grid %>%
  collect_metrics()
```


```{r}
# Visualization is often more helpful to understand model performance.

lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_line(size = 1.5, show.legend = FALSE) +
  facet_wrap(~.metric) +
  scale_x_log10()
```

## Choose the final model based on AUC. 

```{r}
best_auc <- lasso_grid %>%
  select_best("roc_auc")

best_auc
```

```{r}
final_lasso <- finalize_workflow(lasso_wf, best_auc)

final_lasso
```


```{r}
library(vip)

final_lasso %>%
  fit(model_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = best_auc$penalty) %>%
  group_by(Sign) %>%
  top_n(20, wt = abs(Importance)) %>%
  ungroup() %>%
  mutate(Sign = factor(Sign, c("POS", "NEG"), c("Government", "NSAG"))) %>%
  mutate(
    Importance = abs(Importance),
    Variable = str_remove(Variable, "tfidf_notes_"),
    Variable = str_remove(Variable, "admin2_"),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~Sign, scales = "free_y") +
  labs(y = NULL)
```

```{r}
model_final <- last_fit(final_lasso, model_split)

model_final %>%
  collect_metrics()
```

```{r}
model_final %>%
  collect_predictions() %>%
  conf_mat(sub_event_type, .pred_class)
```

